Adversarial Training against Systematic Uncertainty

### 2016-09-07 tex/paper.tex
added toy example without continuous case

\usepackage[final]{nips_2016}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{graphicx}
\usepackage{color}
\usepackage{algorithm}
\usepackage{algpseudocode}

@@

In this setting, the nuisance parameter $Z$ is considered as a random variable,
for which we require the uniform prior $p(z)$ (for $z$ in ${\cal Z}$).
Importantly, classification of $Y$ with respect to $X$ is therefore considered
in the context where $Z$ is marginalized out, which means that the classifier
minimizing ${\cal L}_f$ is optimal with respect to $Y|X$, but not necessarily
with $Y|X,Z$. Results
hold for a nuisance parameter $Z$ taking either categorical values or continuous
values within a bounded support. By abuse of notation, $H(p_Z)$ denotes the
differential entropy in this latter case. Finally, propositions below are derived in a
non-parametric setting, by assuming that both $f$ and $r$ have enough capacity.

@@

Proposition~\ref{prop:2} suggests that if at each step of
Algorithm~\ref{alg:adversarial-training} the adversary $r$ is allowed to reach
its optimum given $f$ (e.g., by setting $K$ sufficiently high) and if $f$ is
updated to improve ${\cal L}_f(\theta_f) - H(p_{Z|f(X;\theta_f)})$, then $f$
should converge to a classifier which is both optimal and pivotal, provided such
a classifier exists. On many problems of interest though, such a classifier may
not exist because the nuisance parameter directly shapes the decision
boundary, in which cases the lower bound $H(p_{Y|X}) - H(p_Z) < {\cal
L}_f(\theta_f) - H(p_{Z|f(X;\theta_f)})$ is strict: $f$ can either be an optimal
classifier or a pivotal quantity, but not both simultaneously. Accordingly the
value function $E$ can be rewritten as
\begin{equation}
    E_\lambda(\theta_f, \theta_r) = {\cal L}_f(\theta_f) - \lambda {\cal L}_r(\theta_f, \theta_r),
\end{equation}
where $\lambda \geq 0$ is a hyper-parameter controlling the trade-off between the performance of $f$
and its independence with respect to the nuisance parameter. Setting $\lambda$ to a large
value will preferrably enforces $f$ to be pivotal while setting $\lambda$ close to $0$
will rather constraint $f$ to be optimal.

>>>>

\usepackage[final]{nips_2016}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{graphicx}
\usepackage{color}
\usepackage{algorithm}
\usepackage{algpseudocode}

@@

In this setting, the nuisance parameter $Z$ is considered as a random variable of bounded support,
for which we require the uniform prior $p(z)$ (for $z \in {\cal Z}$).
Importantly, classification of $Y$ with respect to $X$ is therefore considered
in the context where $Z$ is marginalized out, which means that the classifier
minimizing ${\cal L}_f$ is optimal with respect to $Y|X$, but not necessarily
with $Y|X,Z$. Results hold for a nuisance parameter $Z$ taking either
categorical values or continuous values within a bounded support. By abuse of
notation, $H(p_Z)$ denotes the differential entropy in this latter case.
Finally, propositions below are derived in a non-parametric setting, by assuming
that both $f$ and $r$ have enough capacity.

@@

Proposition~\ref{prop:2} suggests that if at each step of
Algorithm~\ref{alg:adversarial-training} the adversary $r$ is allowed to reach
its optimum given $f$ (e.g., by setting $K$ sufficiently high) and if $f$ is
updated to improve ${\cal L}_f(\theta_f) - H(p_{Z|f(X;\theta_f)})$, then $f$
should converge to a classifier which is both optimal and pivotal, provided such
a classifier exists. On many problems of interest though, such a classifier may
not exist because the nuisance parameter directly shapes the decision
boundary, in which cases the lower bound $H(p_{Y|X}) - H(p_Z) < {\cal
L}_f(\theta_f) - H(p_{Z|f(X;\theta_f)})$ is strict: $f$ can either be an optimal
classifier or a pivotal quantity, but not both simultaneously. In this situation, it is natural to rewrite the
value function $E$  as
\begin{equation}
    E_\lambda(\theta_f, \theta_r) = {\cal L}_f(\theta_f) - \lambda {\cal L}_r(\theta_f, \theta_r),
\end{equation}
where $\lambda \geq 0$ is a hyper-parameter controlling the trade-off between the performance of $f$
and its independence with respect to the nuisance parameter. Setting $\lambda$ to a large
value will preferrably enforces $f$ to be pivotal while setting $\lambda$ close to $0$
will rather constraint $f$ to be optimal.
